---
title: "Main"
author: "vivek gidla"
date: "June 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load csv
```{r}
sp500_training_complete<-read.csv("/home/vivek/Documents/sandbox/SandPTrial/S-PTrial/GSPC.csv")
```

# init plot
```{r}
plot(sp500_training_complete$Close,type="l", col="blue", lwd=2, xlab = "Date", ylab = "Close", main = "SP500 Index Price")
```


Function for minmax scaling source: https://gist.github.com/swayson/b5a6d3cd796ab1d08df1
```{r}
minmax_scaler <- function(x, a, b) {
    "
    x: data. numeric vector of values to be scaled
    a: desired minimum after scaling takes place
    b: desired maximum after scaling takes place
    
    e.g. f(c(1,2,3,4), 1, 17)
    [1]  1.000000  6.333333 11.666667 17.000000
    "
    (((b - a)*(x - min(x))) / (max(x) - min(x))) + a
}
```


Function to invert the scaling
```{r}
inverse_minmax_scaler <- function(x){
  maxval = max(sp500_training_complete$Close) # this is important to do!
  minval = min(sp500_training_complete$Close) # same for this, min max should be of the OG dataset.
  x*(maxval-minval) + minval
}
```

Perform minmax scaling on "Close" values
```{r}
sp500_training_scaled = array(minmax_scaler(sp500_training_complete$Close,0,1))
```

Some global variables
```{r}
N <- 50L #window size for LSTM
total <- NROW(sp500_training_complete) #size of training set
```

Create feature set and labels 
```{r}
modified_sp500_training_scaled = c(sp500_training_scaled, replicate(N, tail(sp500_training_scaled, 1))) #duplicates last row
feats = NULL
labels = NULL
for(i in 1:(total-N))
{
  s = i-1+N
  feats = rbind(feats,modified_sp500_training_scaled[i:s])
  labels = rbind(labels,modified_sp500_training_scaled[s+1])
}
feats = array(feats, dim=c((total-N), N,1))
```

Import Keras
```{r}
library(keras)
use_condaenv("r-tensorflow") # turns out I have to do this.
#install_keras()
```

```{r}
model <- keras_model_sequential()
model %>%
  layer_lstm(units = 50, return_sequences = TRUE, input_shape=c(N, 1),unroll=FALSE) %>%
  layer_dropout(0.2) %>%
  layer_lstm(units = 50, return_sequences = TRUE) %>%
  layer_dropout(0.2) %>%
  layer_lstm(units = 50, return_sequences = TRUE) %>%
  layer_dropout(0.2) %>%
  layer_lstm(units = 50) %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 1)
  
```

Model overview
```{r}
summary(model)
```


```{r}
model %>%
  compile(optimizer = 'adam', loss = 'mean_squared_error')
```

Train the model
```{r}
history = model %>%
  fit(feats, labels, epochs = 100, batch_size = 32)
```


For plotting loss
```{r}
loss = history$metrics$loss
epochs = 1:length(loss)+1
```

```{r}
plot(epochs,loss, title('Training loss'))
```



Prepre test set
```{r}
sp500_total = rbind(sp500_training_complete,sp500_training_complete) #concatenate(duplicate) datasets, one on top of other
sp500_test = sp500_total$Close[(dim(sp500_total)[1]-total-N+1):dim(sp500_total)[1]]
# perform minmax scaling
sp500_testing_scaled = array(minmax_scaler(sp500_test,0,1))
```
As done for the training set, prepare test set of feats (no labels of course)
```{r}
modified_sp500_testing_scaled = c(sp500_testing_scaled, replicate(N, tail(sp500_testing_scaled, 1)))
test_feats = NULL
for(i in 1:total)
{
  s = i-1+N
  test_feats = rbind(test_feats,modified_sp500_testing_scaled[i:s])
}
test_feats = array(test_feats, dim=c((total), N,1))
```

Predict
```{r}
predictions = model %>% predict(test_feats)
```
Invert the scaling 
```{r}
predictions_rescaled = inverse_minmax_scaler(predictions)
```


Plot actual vs predicted
```{r}
plot(predictions_rescaled, type="l", col="red", lwd=2, ylim = c(1400,2900), xlab = "Date", ylab = "Close", main = "SP500 Index Prediction")
lines(sp500_training_complete$Close,type="l",col="blue", lwd=2)
legend(10,2900,legend = c("Predicted","Actual"), fill=c("red","blue"))
```

```{r}
library(Metrics)
```

Calculate Root Mean Squared
```{r}
sqrt(mse(sp500_training_complete$Close[51:322],predictions_rescaled[51:322]))
```

```{r}
typeof(predictions_rescaled)
typeof(sp500_training_complete$Close)
length(sp500_training_complete$Close)

# get direction of actual
(sp500_training_complete$Close[2:length(sp500_training_complete$Close)] - sp500_training_complete$Close[1:(length(sp500_training_complete$Close)-1)]) >= 0

# get direction of predicted
(predictions_rescaled[2:length(predictions_rescaled)] - predictions_rescaled[1:(length(predictions_rescaled)-1)]) >= 0
```



































